---
title: "Mobility in Germany in Times of Coronavirus"
author: "Ansgar Wolsing"
date: "30 4 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vroom)
library(patchwork)

```

## Community Mobility Reports

Community Mobility Reports are available from https://www.google.com/covid19/mobility/

> These datasets show how visits and length of stay at different places change compared to a baseline. We calculate these changes using the same kind of aggregated and anonymized data used to show popular times for places in Google Maps.
> 
> Changes for each day are compared to a baseline value for that day of the week:
> 
> The baseline is the median value, for the corresponding day of the week, during the 5-week period Jan 3–Feb 6, 2020.
The datasets show trends over several months with the most recent data representing approximately 2-3 days ago—this is how long it takes to produce the datasets.


```{r load_data, echo=FALSE, cache=TRUE}

# download CSV from Google
csv_url <- "https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv"
mob <- vroom(csv_url)

```


```{r transform_data, cache=TRUE}

var_suffix <- "_percent_change_from_baseline"
mob_long <- mob %>% 
  select(-country_region_code, -sub_region_2) %>% 
  # transform data into tidy format
  pivot_longer(cols = ends_with(var_suffix), names_to = "activity", values_to = "change") %>%
  # clean activity value strings
  mutate(activity = str_replace(activity, var_suffix, ""),
         sub_region_1 = ifelse(is.na(sub_region_1), "country-wide", sub_region_1))


# limit dataset to Germany
mob_ger_long <- mob_long %>% 
  filter(country_region == "Germany")


```


The dataset contains information on federal state-level and for Germany as a whole. In the latter case, `sub_region_1` is `NA`.


```{r message=FALSE}
plot_region <- function(x, country = "Germany", filter_activity = "any", add_smooth = TRUE, df = mob_ger_long) {
  df <- df %>% 
    filter(country_region %in% country) %>% 
    filter(sub_region_1 %in% x) %>% 
    mutate(activity = str_replace_all(activity, "_", " ") %>% str_to_title())
  
  if (filter_activity != "any") {
    df <- df %>% 
      filter(activity %in% filter_activity)
  }
  p <- df %>% 
    ggplot(aes(date, change)) +
    geom_area(alpha = 0.1, fill = "grey50") +
    labs(x = NULL, y = NULL) +
    theme_minimal()
  if (add_smooth){
    p <- p +
      geom_smooth(se = FALSE, method = "loess", span = 0.25, col = "darkgreen")
  }
  if (length(country) > 1) {
    p <- p + facet_grid(activity ~ country_region)
  } else if (length(filter_activity) > 1 | filter_activity == "any") {
     p <- p + facet_grid(activity ~ sub_region_1)
  } else if (filter_activity == "any") {
    p <- p + facet_wrap(vars(activity))
  } else if (length(x) > 1) {
    p <- p + facet_wrap(vars(sub_region_1), ncol = 1)
  } 
  p
}



```

```{r}

ger_states <- distinct(mob_ger_long, sub_region_1) %>% pull(sub_region_1)
#map(ger_states, plot_region, filter_activity = "any")

```


```{r message=FALSE, fig.height=9}

plot_region("country-wide", filter_activity = "any")

```


```{r message=FALSE, fig.height=15}

#plot_region(c("Bavaria", "North Rhine-Westphalia"), c("workplaces", "retail_and_recreation", "residential"))
plot_region(c("Bavaria", "North Rhine-Westphalia"), filter_activity = "any")


```


```{r message=FALSE, fig.height=15}

plot_region("country-wide", country = c("Germany", "Italy", "United States"), filter_activity = "any",  df = mob_long)

```

